---
title: "SLiPBD Paper Results"
author: "Denise Jennings"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)

source("00_paper_packages.R")
source("00_paper_read_in_datasets.R")
source("01_paper_setup.R")
source("03_infant_death_check.R")
```

## Results 

Following the most recent update, the cohort contained `r nrow(slipbd_database)` babies (fetuses and births) from `r length(unique(slipbd_database$pregnancy_id))` completed or ongoing pregnancies to `r length(unique(slipbd_database$mother_upi))` women. 

Estimated dates of conception ranged from `r min(slipbd_database$est_date_conception)` to `r max(slipbd_database$est_date_conception)`, and end of pregnancy dates for the subset of babies from completed pregnancies ranged from `r min(slipbd_database$date_end_pregnancy, na.rm = T)` to `r max(slipbd_database$date_end_pregnancy, na.rm = T)`. 

<br>

<br>

### **Table 3: Pregnancy outcome for all babies in the SLiPBD cohort, Scotland 2000-Dec 2023** 

```{r pregnancy outcomes table}
# Reorder fetus outcome 1 categories in line with paper format
slipbd_database <- slipbd_database %>% 
  mutate(fetus_outcome1 = factor(fetus_outcome1, levels = c("Live birth", "Stillbirth", "Termination", 
                                                            "Ectopic pregnancy","Molar pregnancy", "Miscarriage",
                                                            "Maternal death", "Unknown", "Unknown - emigrated",
                                                         "Unknown - assumed early loss", "Ongoing")))

# Count of number of fetus/babies for each outcome  
table_3 <- slipbd_database %>%
  group_by(fetus_outcome1) %>% 
  summarise(N = n()) %>% 
  ungroup() %>% 
  
  # Calcuate percentages based on total cohort number
  mutate(total_pregnancies = nrow(slipbd_database),
         outcome_percent = N/total_pregnancies*100, 
         outcome_percent = round(outcome_percent, 2)) %>% 
  
  # Drop total column
  select(-total_pregnancies) %>% 
  mutate(outcome_percent = paste0("(", outcome_percent, ")")) %>% 
  
  # Combine N and percent in one column
  unite("Number_(%)_of_babies", N, outcome_percent, sep = " ") 
  
  # Present data in table
  table_3 %>% 
  kable(col.names = c("Pregnancy outcome", "Number (%) of babies"))


```

<br>

<br>

### **Figure 2: Number of babies in the SLiPBD cohort, by pregnancy outcome group and year of conception, Scotland 2000-Dec 2023**
```{r pregnancy outcomes chart}

# Assign fetus outcomes into key outcome group categories
slipbd_database <- slipbd_database %>% 
  mutate(key_outcome_groups = case_when(fetus_outcome1 == "Miscarriage" ~ "Spontaneous early pregnancy loss",
                                        fetus_outcome1 == "Ectopic pregnancy" ~ "Spontaneous early pregnancy loss",
                                        fetus_outcome1 == "Molar pregnancy" ~ "Spontaneous early pregnancy loss",
                                        fetus_outcome1 == "Unknown" ~ "Unknown and maternal death",
                                        fetus_outcome1 == "Unknown - assumed early loss" ~ "Unknown and maternal death",
                                        fetus_outcome1 == "Unknown - emigrated" ~ "Unknown and maternal death",
                                        fetus_outcome1 == "Maternal death" ~ "Unknown and maternal death", 
                                        TRUE ~ fetus_outcome1),
         # Reorder the levels
         key_outcome_groups = factor(key_outcome_groups, levels = c("Live birth", "Stillbirth", "Termination", 
                                                                    "Spontaneous early pregnancy loss",
                                                                    "Unknown and maternal death", "Ongoing")))

# Then present number of babies in the SLiPBD cohort, by pregnancy outcome group and year of conception 
figure_2 <- slipbd_database %>% 
  mutate(conceptionyear = factor(conceptionyear)) %>% 
  ggplot(aes(x = conceptionyear, fill = key_outcome_groups)) +
  geom_bar() +
  scale_x_discrete(breaks = every_nth(n = 2)) +
  xlab("Year of conception") +
  ylab("Number of babies") +
  labs(fill = "Pregnancy outcomes") +
  theme(axis.text.x = element_text(size=8)) +
  scale_fill_manual(values = phs_colours(c("phs-purple", "phs-magenta", "phs-blue", "phs-green", "phs-graphite", "phs-teal")))  

figure_2 + theme

ggsave("figure_2.png", dpi = 300)

```

<br>

<br>

### **Figure 3: Gestation at end of pregnancy for babies in the SLiPBD cohort with known pregnancy outcome, Scotland 2000-Dec 2023** 
```{r}
# Subset data to count number of weeks gestation for pregnancy outcomes
gest_count <- slipbd_database %>% 
  group_by(fetus_outcome1, gest_end_pregnancy) %>% 
  summarise(N = n()) %>% 
  ungroup()
gest_count <- gest_count %>% mutate(panel_no = case_when(fetus_outcome1 =="Live birth" ~ "a)",
                                                         T~"b)"))
# Plot number of weeks gestation at end of pregnancy by outcome 
# For ease of viewing, create 2 charts, one with live births, the other with remaining outcomes

##use facet_wrap - need to add a and b labels
p <- gest_count %>%
   filter(fetus_outcome1 %in% c("Live birth", "Stillbirth", "Termination", 
                          "Ectopic pregnancy", "Molar pregnancy", "Miscarriage")) %>% 
  ggplot(aes(x = gest_end_pregnancy, y = N, group=fetus_outcome1)) +
  geom_line(aes(colour = fetus_outcome1), lwd = 1) +
  facet_wrap(vars(panel_no), scales = "free_y")+
  xlab("Number of weeks") +
  ylab("Number of babies") +
  #xlim(4, 44) +
  scale_x_continuous(limit = c(4,44), 
                         breaks = c(4,8,12, 16, 20, 24, 28, 32, 36, 40, 44))+
  labs(colour = "Pregnancy outcome") +
  theme_bw() +
  theme(text = element_text(size=12),
    legend.position = "bottom",
        legend.justification = "left", 
         strip.text = element_text(hjust = 0, size=16, face="bold"),
         strip.background = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(colour = "black")
        ) +
  theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size=10), 
        legend.justification = "left") +
  scale_colour_manual(values = 
                        phs_colours(c("phs-purple", "phs-magenta", "phs-blue",
                                      "phs-green", "phs-rust", "phs-teal")))  

p


ggsave("figure_3_wrap.png", units = "in", width = 9, height  = 4.5, dpi = 300)



rm(gest_count)

```

<br>

<br>

### **Table 4: Source of information on gestation at end of pregnancy for babies in the SLiPBD cohort with known pregnancy outcome, Scotland 2000-Dec 2023**

```{r}
# Subset a count of known pregnancy outcomes
known_outcome_totals <- slipbd_database %>% 
  filter(fetus_outcome1 %in% c("Live birth", "Stillbirth", "Termination", 
                               "Ectopic pregnancy", "Molar pregnancy", "Miscarriage")) %>% 
  group_by(fetus_outcome1) %>% 
  summarise(N = n())

# Crosstab table of pregnancy outcome and gestation ascertainment
table_4 <- slipbd_database %>% 
  filter(fetus_outcome1 %in% c("Live birth", "Stillbirth", "Termination", 
                               "Ectopic pregnancy", "Molar pregnancy", "Miscarriage")) %>% 
  group_by(fetus_outcome1, gestation_ascertainment) %>% 
  summarise(N = n()) %>% 
  spread(gestation_ascertainment, N) %>% 
  ungroup() %>% 
  
  # Rename columns
  rename(recorded = `Gestation recorded on end of pregnancy record`,
         calculated = `Gestation calculated from gestation at booking`,
         imputed = `Gestation imputed based on outcome of pregnancy`) %>% 
  
  # Add in total column from known outcome totals to calculate percentage for each value
  mutate(total_number_of_babies = known_outcome_totals$N) %>% 
  
  # Reorder columns
  select(fetus_outcome1, total_number_of_babies, recorded, calculated, imputed) %>% 
  
  # Add total row
  adorn_totals("row") %>% 
  
  # Calculate percentages
  mutate(recorded_percent = recorded/total_number_of_babies*100,
         calculated_percent = calculated/total_number_of_babies*100,
         imputed_percent = imputed/total_number_of_babies*100,
         recorded_percent = round(recorded_percent, 2),
         calculated_percent = round(calculated_percent, 2),
         imputed_percent = round(imputed_percent, 2)) %>% 

  
# Combine numbers and percentages
    mutate(recorded_percent = paste0("(", recorded_percent, ")"),
           calculated_percent = paste0("(", calculated_percent, ")"),
           imputed_percent = paste0("(", imputed_percent, ")")) %>% 
    unite("Recorded", recorded, recorded_percent, sep = " ") %>% 
    unite("Calculated", calculated, calculated_percent, sep = " ") %>% 
    unite("Imputed", imputed, imputed_percent, sep = " ") 

table_4 %>% 
  kable(col.names= c("Pregnancy outcome", "Total number of babies", "Available on end of pregnancy record", "Calculated from gestation on antenatal booking record", "Imputed based on pregnancy outcome type"))  

 rm(known_outcome_totals)
 
```

<br>

<br>

### **Table 5: Key maternal sociodemographic characteristics for all babies in the SLiPBD cohort, Scotland 2000-Dec 2023**

```{r maternal age table}

# Create new variable with age categories
slipbd_database <- slipbd_database %>%
  mutate(mat_age_cat = case_when(between(maternal_age_conception, 0, 19) ~ "<20",
                                 between(maternal_age_conception, 20, 24) ~ "20-24",
                                 between(maternal_age_conception, 25, 29) ~ "25-29",
                                 between(maternal_age_conception, 30, 34) ~ "30-34",
                                 between(maternal_age_conception, 35, 39) ~ "35-39",
                                 maternal_age_conception >= 40  ~ "≥40",
                                 is.na(maternal_age_conception) ~ "Unknown/missing"))


levels_order <- c("<20", "20-24", "25-29", "30-34", "35-39", "≥40", "Unknown/missing")

slipbd_database$mat_age_cat <- factor(slipbd_database$mat_age_cat, levels = levels_order)


# Function to count and add percentage columns for key maternal sociodemographics
sociodemographic_counts <- function(data, variable){
  
  data %>% 
  group_by({{variable}}) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(percent = n/nrow(data)*100,
         percent = round(percent, 2))
}

# Count of fetus/babies for each age category
maternal_age <- sociodemographic_counts(slipbd_database, mat_age_cat)

# Present in table
maternal_age %>% 
  kable(col.names = c("Maternal age at conception (years)", "Number of babies", "%"))

# Calculation of overall median (min-max) maternal age at conception
median_table <- slipbd_database %>% 
  mutate(median = median(maternal_age_conception, na.rm = T),
         min = min(maternal_age_conception, na.rm = T), 
         max = max(maternal_age_conception, na.rm = T)) %>% 
  select(median, min, max) %>% 
  slice(1)

# Meadian (min-max) age table
median_table %>% 
  kable(col.names = c("Median", "Min", "Max"))


# Add count for number of pregnancies for each age category
# Create subset for number of pregnancies including age category variable
pregnancies_cohort <- slipbd_database %>% 
  group_by(pregnancy_id) %>% 
  filter(fetus_marker == 1) %>% 
  ungroup() 

# Count or pregnancies by age category
pregnancies_maternal_age <- sociodemographic_counts(pregnancies_cohort, mat_age_cat)

# Present in table
pregnancies_maternal_age %>% 
  kable(col.names = c("Maternal age at conception (years)", "Number of pregnancies", "%"))

# Calculation of median (min-max) maternal age at conception for number of pregnancies 
pregnancies_median_table <- pregnancies_cohort %>% 
  mutate(median = median(maternal_age_conception, na.rm = T),
         min = min(maternal_age_conception, na.rm = T), 
         max = max(maternal_age_conception, na.rm = T)) %>% 
  select(median, min, max) %>% 
  slice(1)

# Meadian (min-max) age table for number of pregnancies
pregnancies_median_table %>% 
  kable(col.names = c("Median", "Min", "Max"))

```

```{r maternal simd}
# Count of maternal SIMD at booking or at end of pregnancy 
# Create new variable to identify if simd is present at booking or and end of pregnancy
slipbd_database <- slipbd_database %>% 
  mutate(any_simd = case_when(is.na(maternal_SIMD_end_preg) ~ maternal_SIMD_booking,
                              maternal_SIMD_end_preg == "Unknown" ~ maternal_SIMD_booking,
                              TRUE ~ maternal_SIMD_end_preg))

# Count number of fetus/babies and add percentages
maternal_simd <- sociodemographic_counts(slipbd_database, any_simd)

# Present in table
maternal_simd %>% 
  kable(col.names = c("Maternal deprivation level (SIMD quintile)", "Number of babies", "%"))

# Follow above steps to add count for number of pregnancies
pregnancies_cohort <- pregnancies_cohort %>% 
  mutate(any_simd = case_when(is.na(maternal_SIMD_end_preg) ~ maternal_SIMD_booking,
                              maternal_SIMD_end_preg == "Unknown" ~ maternal_SIMD_booking,
                              TRUE ~ maternal_SIMD_end_preg))

pregnancies_maternal_simd <- sociodemographic_counts(pregnancies_cohort, any_simd)

# Pregnancies table
pregnancies_maternal_simd %>% 
  kable(col.names = c("Maternal deprivation level (SIMD quintile)", "Number of pregnanices", "%"))

```


```{r maternal ethnicity table}

# Code all NAs as unknown/missing
slipbd_database <- slipbd_database %>% 
  mutate(maternal_ethnicity = case_when(is.na(maternal_ethnicity) ~ "5 Unknown/missing",
                                        TRUE ~ maternal_ethnicity))
# Count of maternal ethnicity         
maternal_ethnicity <- sociodemographic_counts(slipbd_database, maternal_ethnicity) 

# Maternal ethnicity table
maternal_ethnicity %>% 
  kable(col.names = c("Maternal ethnicity", "Number of babies", "%"))


# Count for number of pregnancies 
# Code all NAs as unknown/missing in number of pregnancies subset
pregnancies_cohort <- pregnancies_cohort %>% 
  mutate(maternal_ethnicity = case_when(is.na(maternal_ethnicity) ~ "5 Unknown/missing",
                                        TRUE ~ maternal_ethnicity))

# Count of maternal ethnicity for number of pregnancies subset       
pregnancies_maternal_ethnicity <- sociodemographic_counts(pregnancies_cohort, maternal_ethnicity) 

pregnancies_maternal_ethnicity %>% 
  kable(col.names = c("Maternal ethnicity", "Number of pregnancies", "%"))

```

<br>

<br>

##  N babies and (separately) pregnancies with/without mat ethnicity. Conceptions from 1 Jan 2022 onwards
```{r}
# Breakdown of maternal ethnicity (just available/missing) for all babies with date of conception from 2022 onwards by pregnancy outcome.

# Create subset of babies from 2022 onwards and ethnicity flag variable
ethnicity <- slipbd_database %>% 
  filter(year(est_date_conception) >= 2022) %>% 
  mutate(ethnicity_flag = case_when(is.na(maternal_ethnicity) ~ "Missing", 
                                    maternal_ethnicity == "5 Unknown/missing" ~ "Missing",
                                    TRUE ~ "Available")) 
#min(ethnicity$est_date_conception)
#table(ethnicity$ethnicity_flag)

# Total count of outcomes from 2022 onwards (to calculate percentage below)
outcome_totals_ethnicity_cohort <- ethnicity %>% 
  group_by(key_outcome_groups) %>% 
  summarise(total = n()) %>% 
  ungroup()

# Count of ethnicity flag by key outcome variables
ethnicity_by_outcome <- ethnicity %>% 
  group_by(key_outcome_groups, ethnicity_flag) %>% 
  summarise(N = n()) %>% 
  ungroup()

# Join with full outcome count
ethnicity_by_outcome <- full_join(ethnicity_by_outcome, outcome_totals_ethnicity_cohort)

# Calculate percent of outcomes with ethnicity available/missing
ethnicity_by_outcome <- ethnicity_by_outcome %>% 
  mutate(percent = N/total*100,
         percent = round(percent, 2))

ethnicity_by_outcome %>% 
  select(-total) %>% 
  kable(col.names = c("Pregnancy outcome", "Ethnicity status", "N of babies", "%"))

# Repeat above steps to add count for number of pregnancies
# Use number of pregnancies subset, filter for 2022 onwards and create ethnicity flag
pregnancies_ethnicity <- pregnancies_cohort %>% 
  filter(year(est_date_conception) >= 2022) %>% 
  mutate(ethnicity_flag = case_when(is.na(maternal_ethnicity) ~ "Missing", 
                                    maternal_ethnicity == "5 Unknown/missing" ~ "Missing",
                                    TRUE ~ "Available"))

# Total count of outcomes from 2022 onwards (to calcualte percentage below)
pregnancies_outcome_totals <- pregnancies_ethnicity %>% 
  group_by(key_outcome_groups) %>% 
  summarise(total = n()) %>% 
  ungroup()

# Count of ethnicity flag by key outcome variables
pregnancies_ethnicity_by_outcome <- pregnancies_ethnicity %>% 
  group_by(key_outcome_groups, ethnicity_flag) %>% 
  summarise(N = n()) %>% 
  ungroup()

# Join with full outcome count
pregnancies_ethnicity_by_outcome <- full_join(pregnancies_ethnicity_by_outcome, pregnancies_outcome_totals)

# Calculate percent of outcomes with ethnicity available/missing
pregnancies_ethnicity_by_outcome <- pregnancies_ethnicity_by_outcome %>% 
  mutate(percent = N/total*100,
         percent = round(percent, 2))

pregnancies_ethnicity_by_outcome %>% 
  select(-total) %>% 
  kable(col.names = c("Pregnancy outcome", "Ethnicity status", "N of pregnancies", "%"))

rm(outcome_totals_ethnicity_cohort, pregnancies_outcome_totals)

```

<br>

<br> 

### **Figure 4: Distribution of maternal age at conception for all babies in the SLiPBD cohort, Scotland 2000-Dec 2023**

```{r maternal age chart}
figure_4 <- slipbd_database %>% 
  ggplot(aes(x = maternal_age_conception)) +
  geom_histogram(fill = "#3F3685", col = "black", binwidth = 1) +
  xlab("Maternal age at conception") +
  ylab("Number of babies")

figure_4 + theme

ggsave("figure_4.png", dpi = 300)
```


<br>

<br>

### **Table 6: Key infant sociodemographic characteristics for live births in the SLiPBD cohort, Scotland 2000-Dec 2023**

```{r baby sex}

# Count of baby sex for live births with percentage column
baby_sex <- slipbd_database %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  group_by(baby_sex) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(percent = n/sum(n)*100,
         percent = round(percent, 2))

baby_sex%>% 
  kable(col.names = c("Baby sex", "Number of live births", "%"))

```

```{r baby ethnicity}
# Count of baby ethnicity for live births with percentage column
baby_ethnicity <- slipbd_database %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  group_by(baby_ethnicity) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(percent = n/sum(n)*100,
         percent = round(percent, 2))

baby_ethnicity %>% 
  kable(col.names = c("Baby ethnicity", "Number of live births", "%"))
```

<br>

<br>

### Chart of completness for baby ethnicity by year of conception (live births only)

```{r}
# Chart of completeness of baby ethnicity by year of conception for live births only . 
total_records <- slipbd_database %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  group_by(conceptionyear) %>% 
  summarise(total_records_per_year = n()) %>% 
  ungroup()

baby_ethnicity_year <- slipbd_database %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  mutate(baby_ethnicity_flag = case_when(is.na(baby_ethnicity) ~ "missing", 
                                    baby_ethnicity == "5 Unknown/missing" ~ "missing",
                                    TRUE ~ "available")) %>% 
  filter(baby_ethnicity_flag == "available") %>% 
  group_by(conceptionyear, baby_ethnicity_flag) %>% 
  summarise(n = n()) %>% 
  ungroup() 

available_baby_ethnicity <- full_join(total_records, baby_ethnicity_year)

baby_ethnicity_completeness <- available_baby_ethnicity %>% 
  #filter(baby_ethnicity_flag == "available") %>% 
  ggplot(aes(x = factor(conceptionyear), y = n/total_records_per_year*100, group = baby_ethnicity_flag)) +
  geom_line(aes(colour = "Ethnicity recorded")) +
  xlab("Year of conception") +
  ylab("% of live births with baby ethniciy available") +
  ylim(0, 100) +
  scale_x_discrete(breaks = every_nth(n = 2)) +
  theme(legend.title = element_blank())

baby_ethnicity_completeness + theme

ggsave("baby_ethnicity_completeness.png", dpi = 300)

rm(total_records, baby_ethnicity_year, available_baby_ethnicity)

```


N babies with/without baby ethnicity born from 1 Jan 2022 onwards
```{r}
# Breakdown of baby ethnicity (just available/missing) for live births with date of conception from 2022 onwards.
# Create baby ethnicity flag
baby_ethnicity <- ethnicity %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  mutate(baby_ethnicity_flag = case_when(baby_ethnicity == "5 Unknown/missing" ~ "Missing",
                                    TRUE ~ "Available"))

# Count number of live births with ethnicity and add percentage  
count_baby_ethnicity_flag <- baby_ethnicity %>% 
  group_by(baby_ethnicity_flag) %>% 
  summarise(N = n()) %>% 
  ungroup() %>% 
  mutate(percent = N/nrow(baby_ethnicity)*100,
         percent = round(percent, 2))

count_baby_ethnicity_flag %>% 
  kable(col.names = c("Ethnicity status", "N of babies", "%"))


# Repeat above steps to count number of pregnancies
# Use number of pregnancies subset, filter and create new baby ethnicity flag variable
pregnancies_baby_ethnicity <- pregnancies_ethnicity %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
   mutate(baby_ethnicity_flag = case_when(baby_ethnicity == "5 Unknown/missing" ~ "Missing",
                                    TRUE ~ "Available"))

# Count number of pregnancies (live births only) with ethnicity and add percentage 
count_pregnancies_baby_ethnicity <- pregnancies_baby_ethnicity %>% 
  group_by(baby_ethnicity_flag) %>% 
  summarise(N = n()) %>% 
  ungroup() %>% 
    mutate(percent = N/nrow(pregnancies_baby_ethnicity)*100,
         percent = round(percent, 2))


count_pregnancies_baby_ethnicity %>% 
  kable(col.names = c("Ethnicity status", "N of pregnancies", "%"))

rm(baby_ethnicity, pregnancies_baby_ethnicity)

```


```{r gestation table}

# Subset for live births and create variable for gestation categories 
gestation_live_births <- slipbd_database %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  mutate(gestation_categories = case_when(gest_end_pregnancy <32 ~ "Very preterm (<32)",
                                          between(gest_end_pregnancy, 32, 36) ~ "Moderately preterm (32-36)",
                                          between(gest_end_pregnancy, 37, 41) ~ "Term (37-41)",
                                          gest_end_pregnancy >= 42 ~ "Post-term (≥42)",
                                          TRUE ~ "Unknown/missing"),
         gestation_categories = factor(gestation_categories, levels = c("Very preterm (<32)", "Moderately preterm (32-36)", 
                                                                        "Term (37-41)", "Post-term (≥42)",
                                                                        "Unknown/missing")))


# Count total number of babies per categories and add percentage column
gestation_categories_count <- gestation_live_births %>% 
  group_by(gestation_categories) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(percent = n/sum(n)*100,
         percent = round(percent, 2)) 

gestation_categories_count %>% 
  kable(col.names = c("Gestation at birth (completed weeks)", "N of all LBs", "%"))

# Calculate median, min and max values for whole cohort of live births
gestation_median_min_max <- gestation_live_births %>% 
   mutate(median = median(gest_end_pregnancy),
         min = min(gest_end_pregnancy), 
         max = max(gest_end_pregnancy)) %>% 
  select(median, min, max) %>% 
  slice(1)

gestation_median_min_max  %>% 
  kable(col.names = c("Median", "Min", "Max"))


# Repeat steps for subset splitting births into single or multiple pregnancies
# Create plurality variable 
gestation_plurality <- gestation_live_births %>% 
  mutate(plurality = if_else(number_of_babies == 1, "Singleton", "Multiple"))
  
single_multiple_gestation <- gestation_plurality %>% 
group_by(plurality, gestation_categories) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = plurality,
              values_from = n) %>% 
  mutate(percent_single = Singleton/sum(Singleton)*100,
         percent_multiple = Multiple/sum(Multiple)*100,
         percent_single = round(percent_single, 2),
         percent_multiple = round(percent_multiple, 2)) %>% 
  select(gestation_categories, Singleton, percent_single, Multiple, percent_multiple)

single_multiple_gestation %>% 
  kable(col.names = c("Gestation at birth (completed weeks)", "N of LB Singleton", "% of LB Singleton", "N of LB Multiple", "% of LB Multiple"))

# Subset for singletons and calculate median, min and max values
singleton_gestation_median_min_max <- gestation_plurality %>% 
  filter(plurality == "Singleton") %>% 
   mutate(median = median(gest_end_pregnancy),
         min = min(gest_end_pregnancy), 
         max = max(gest_end_pregnancy)) %>% 
  select(median, min, max) %>% 
  slice(1)

singleton_gestation_median_min_max  %>% 
  kable(col.names = c("Singleton Median gestation", "Min", "Max"))

# Subset for multiples and calculate median, min and max values
multiple_gestation_median_min_max <- gestation_plurality %>% 
  filter(plurality == "Multiple") %>% 
   mutate(median = median(gest_end_pregnancy),
         min = min(gest_end_pregnancy), 
         max = max(gest_end_pregnancy)) %>% 
  select(median, min, max) %>% 
  slice(1)

multiple_gestation_median_min_max  %>% 
  kable(col.names = c("Multiple Median gestation", "Min", "Max"))

rm(gestation_live_births, gestation_plurality)

```

```{r birthweight table}

# Filter for live births and create variable for birthweight categories (removing unrealistic weights by categorising them as Unknown/missing)
birthweight_live_births <- slipbd_database %>% 
  filter(fetus_outcome1 == "Live birth") %>% 
  mutate(birthweight = case_when(birthweight <100 ~ NA, 
                                 birthweight >7999 ~ NA,
                                 T ~ birthweight),
         birthweight_cat = case_when(between(birthweight, 100, 1499) ~ "Very low (<1500)",
                                       between(birthweight, 1500, 2499) ~ "Low (1500-2499)",
                                       between(birthweight, 2500, 3999) ~ "Typical (2500-2499)",
                                       between(birthweight, 4000, 7999) ~ "High (≥4000)",
                                       TRUE ~ "Unknown/missing"), 
        birthweight_cat = factor(birthweight_cat, levels = c("Very low (<1500)", "Low (1500-2499)", "Typical (2500-2499)", "High (≥4000)", "Unknown/missing")))


# Count total number of babies per birthweight category and add percentage column (whole cohort of live births)
birthweight <- birthweight_live_births %>% 
  group_by(birthweight_cat) %>% 
  summarise(N = n()) %>% 
  ungroup() %>% 
  mutate(percent = N/sum(N)*100,
         percent = round(percent, 2))

birthweight %>% 
  kable(col.names = c("Birthweight", "N of all LBs", "%"))

# Calculate median, min and max values for whole cohort of live births
birthweight_median_min_max <- birthweight_live_births %>% 
   mutate(median = median(birthweight, na.rm = T),
         min = min(birthweight, na.rm = T), 
         max = max(birthweight, na.rm = T)) %>% 
  select(median, min, max) %>% 
  slice(1)

birthweight_median_min_max  %>% 
  kable(col.names = c("Median", "Min", "Max"))

# Create subset for singletons and multiples and count number of babies per birthweight category  
birthweight_plurality <- birthweight_live_births %>% 
  mutate(plurality = if_else(number_of_babies == 1, "Singleton", "Multiple")) 

single_multiple_birthweight <- birthweight_plurality %>% 
  group_by(plurality, birthweight_cat) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  
  # Pivot wider to separate singleton and multiple births
  pivot_wider(names_from = plurality,
              values_from = n) %>% 
  mutate(percent_single = Singleton/sum(Singleton)*100,
         percent_multiple = Multiple/sum(Multiple)*100,
         percent_single = round(percent_single, 2),
         percent_multiple = round(percent_multiple, 2)) %>% 
  select(birthweight_cat, Singleton, percent_single, Multiple, percent_multiple)
    
single_multiple_birthweight %>% 
  kable(col.names = c("Birthweight", "N of LB Singleton", "% of LB Singleton", "N of LB Multiple", "% of LB Multiple"))


# Subset for singletons and calculate median, min and max values
singleton_birthweight_median_min_max <- birthweight_plurality %>% 
  filter(plurality == "Singleton") %>% 
   mutate(median = median(birthweight, na.rm = T),
         min = min(birthweight, na.rm = T), 
         max = max(birthweight, na.rm = T)) %>% 
  select(median, min, max) %>% 
  slice(1)

singleton_birthweight_median_min_max  %>% 
  kable(col.names = c("Singleton Median birthweight", "Min", "Max"))

# Subset for multiples and calculate median, min and max values
multiple_birthweight_median_min_max <- birthweight_plurality %>% 
  filter(plurality == "Multiple") %>% 
   mutate(median = median(birthweight, na.rm = T),
         min = min(birthweight, na.rm = T), 
         max = max(birthweight, na.rm = T)) %>% 
  select(median, min, max) %>% 
  slice(1)

multiple_birthweight_median_min_max  %>% 
  kable(col.names = c("Multiple Median birthweight", "Min", "Max"))

rm(birthweight_live_births, birthweight_plurality)
```

<br>

<br>

## **Table 7: Completeness of valid maternal CHI number for all babies in the SLiPBD cohort by pregnancy outcome, Scotland 2000-Dec 2023** 

```{r}

# Check and count number of valid maternal CHI grouped by outcome
# Count of totals to calculate percent
outcome_totals <- slipbd_database %>% 
  group_by(fetus_outcome1) %>% 
  summarise(N = n())

valid_maternal_chi <- slipbd_database %>% 
  filter(chi_check(mother_upi) == "Valid CHI") %>% 
  group_by(fetus_outcome1) %>% 
  summarise(number_of_valid = n()) %>% 
  ungroup() %>% 
  mutate(percent_of_valid = number_of_valid/outcome_totals$N*100,
         percent_of_valid = round(percent_of_valid, 2)) %>% 
  adorn_totals("row")
    
valid_maternal_chi %>% 
  kable(col.names = c("Pregnancy outcome", "Number (%) with a valid maternal CHI number available", "%"))

rm(outcome_totals)

```

<br>

<br>

## Paper text 
```{r}
# Check number of valid baby CHI for live births
valid_baby_chi <- slipbd_database %>% 
  filter(chi_check(baby_chi) == "Valid CHI") %>% 
  group_by(fetus_outcome1) %>% 
  summarise(number_of_valid = n())
  
valid_baby_chi %>% 
  kable(col.names = c("Pregnancy outcome", "Number of valid CHIs"))

# Check number of linked and dropped SMR02 records for live and stillbirths

# Filter final variables data for live and stillbirths
all_variables_births <- all_variables %>%
  filter(fetus_outcome1 == "Live birth" | fetus_outcome1 == "Stillbirth") 

# Filter SMR02 for live and stillbirths
smr02_births <- smr02_data %>%
  filter(smr02_outcome_type == "Live birth" | smr02_outcome_type == "Stillbirth")

# Find event ids that are in SMR02 and the final variables dataset         
smr02_births_linked <- smr02_births %>%
  filter(event_id %in% all_variables_births$births_smr02_live_births) %>% 
  group_by(smr02_outcome_type) %>% 
  summarise(n = n())

smr02_births_linked %>% 
  kable(col.names = c("Pregnancy outcome", "Linked records"))

# Check number of SMR02 delivery records (live and stillbirths)
nrow(smr02_births)

# Check number of dropped records
nrow(smr02_births) - sum(smr02_births_linked$n)

# Check infant death records for deaths not linked to a corresponding live birth within the SLiPBD. 
not_in_cohort <- not_in_cohort %>% 
  filter(year_of_registration != 1999 | is.na(year_of_registration))
nrow(not_in_cohort)

# Check unlinked for ones with no nrs triplicate id + 1 for NRS triplicate ID that is invalid (B00000)
sum(is.na(not_in_cohort$nrs_triplicate_id)) + 1

rm(all_variables_births) 
  
```

### **Supplementary appendix 2: Number of babies in SLiPBD with known pregnancy outcome compared to number of end of pregnancy events recorded in source records, Scotland 2000-Dec 2023** 
```{r}

# Function to count outcomes in each dataset
single_outcome_count <- function(data, outcome){
  
  data %>% 
    group_by({{outcome}}) %>% 
    summarise(N = n())
}

duel_outcome_count <- function(outcome)
  
  slipbd_database %>% 
  filter(fetus_outcome1 == {{outcome}} | fetus_outcome2 == {{outcome}}) 


# Slipbd outcome count
slipbd_outcome_count <- single_outcome_count(slipbd_database, fetus_outcome1)

slipbd_outcome_count %>% 
  kable(col.names = c("SLiPBD Pregnancy outcome", "N"))


# Live births in NRS
nrslb_count <- single_outcome_count(nrs_live_births, nrslb_outcome_type)

nrslb_count %>% 
  kable(col.names = c("NRS LB Pregnancy outcome", "N"))

lb_duel_outcome_count <- duel_outcome_count("Live birth")
nrow(lb_duel_outcome_count)


# Stillbirths in NRS
nrs_sb_count <- single_outcome_count(nrs_sb, nrssb_outcome_type)

nrs_sb_count %>% 
  kable(col.names = c("NRS SB Pregnancy outcome", "N"))

sb_duel_outcome_count <- duel_outcome_count("Stillbirth")
nrow(sb_duel_outcome_count)



# Terminations in ToPS/AAS
tops_count <- single_outcome_count(tops_aas, aas_outcome_type)

tops_count %>% 
  kable(col.names = c("ToPS Pregnancy outcome", "N"))

#max(tops_aas$aas_date_of_termination)
#max(slipbd_database$date_end_pregnancy, na.rm = T)

# termination in slipbd with duel outcomes
duel_termination_count <- duel_outcome_count("Termination")
nrow(duel_termination_count)


# Outcome count in SMR01
smr01_count <- single_outcome_count(smr01, smr01_outcome_type)

smr01_count %>% 
  kable(col.names = c("SMR01 Pregnancy outcome", "N"))

# Outcome count in SMR02
smr02_nonlive_count <- single_outcome_count(smr02_nonlive, smr02_outcome_type)

smr02_nonlive_count %>% 
  kable(col.names = c("SMR02 Pregnancy outcome", "N"))

# Duel outcome ectopic pregnancies
duel_ectopic_count <- duel_outcome_count("Ectopic pregnancy")
nrow(duel_ectopic_count)

# Duel outcome molar pregnancies
duel_molar_count <- duel_outcome_count("Molar pregnancy")
nrow(duel_molar_count)

# Duel outcome miscarriages
duel_miscarriage_count <- duel_outcome_count("Miscarriage")
nrow(duel_miscarriage_count)

# Minimum dates for all datasets.
min(nrs_live_births$nrslb_date_of_birth)
min(nrs_sb$nrssb_date_of_birth)
min(tops_aas$aas_date_of_termination)
min(smr01$smr01_cis_admission_date)
min(smr02_data$smr02_admission_date)
min(smr02_nonlive$smr02_admission_date)


```






